Actuator line modeling of vertical-axis turbines

### 2016-05-04 paper.tex
Define NREL earlier

In momentum methods, the inflow velocity is determined by solving for the axial
and angular induction factors \cite{Manwell2002}. However, using Navier--Stokes
methods, it is somewhat unclear how to calculate the velocity vector used to
compute the angle of attack and relative velocity, though we have access to much
more information about the flow. Sorensen and Shen used an actuator line
element's position to determine the inflow velocity for an axial-flow turbine
\cite{Sorensen2002}. Similarly, Shamsoddin and Porte-Agel use the velocity at a
blade element's location in their actuator line simulation of a vertical-axis
turbine using LES \cite{Shamsoddin2014}. NREL's SOWFA ALM for OpenFOAM also uses
the velocity at the actuator line element location for computing inflow velocity
and local angle of attack with no corrections \cite{Churchfield2013}.

@@

The US National Renewable Energy Laboratory (NREL) has developed and released an
actuator line modeling library, SOWFA~\cite{Churchfield2014b}, for simulating
horizontal-axis wind turbine arrays using the OpenFOAM finite volume CFD
library. OpenFOAM is free, open-source, widely used throughout industry and
academia, and has grown a very active support community around itself
\cite{Bachant2015-CFD-online}. Though SOWFA is also open-source, and has the
ability to be coupled to NREL's FAST HAWT BEM code, its mostly procedural style
would have required significant effort and duplicate code to adapt for
cross-flow turbines. Thus, a new and more general ALM library was developed from
the ground up that could model both cross- and axial-flow turbines, as well as
standalone actuator lines. The actuator line model developed here, dubbed
turbinesFoam \cite{Bachant2016-turbinesFoam-v0.0.7}, was also written as an
extension library for OpenFOAM, and was developed freely and openly from its
inception in order to increase community engagement and research efficiency.

>>>>

In momentum methods, the inflow velocity is determined by solving for the axial
and angular induction factors \cite{Manwell2002}. However, using Navier--Stokes
methods, it is somewhat unclear how to calculate the velocity vector used to
compute the angle of attack and relative velocity, though we have access to much
more information about the flow. Sorensen and Shen used an actuator line
element's position to determine the inflow velocity for an axial-flow turbine
\cite{Sorensen2002}. Similarly, Shamsoddin and Porte-Agel use the velocity at a
blade element's location in their actuator line simulation of a vertical-axis
turbine using LES \cite{Shamsoddin2014}. The US National Renewable Energy
Laboratory (NREL) SOWFA ALM for OpenFOAM also uses the velocity at the actuator
line element location for computing inflow velocity and local angle of attack
with no corrections \cite{Churchfield2013}.

@@

NREL has developed and released an actuator line modeling library,
SOWFA~\cite{Churchfield2014b}, for simulating horizontal-axis wind turbine
arrays using the OpenFOAM finite volume CFD library. OpenFOAM is free,
open-source, widely used throughout industry and academia, and has grown a very
active support community around itself \cite{Bachant2015-CFD-online}. Though
SOWFA is also open-source, and has the ability to be coupled to NREL's FAST HAWT
BEM code, its mostly procedural style would have required significant effort and
duplicate code to adapt for cross-flow turbines. Thus, a new and more general
ALM library was developed from the ground up that could model both cross- and
axial-flow turbines, as well as standalone actuator lines. The actuator line
model developed here, dubbed turbinesFoam
\cite{Bachant2016-turbinesFoam-v0.0.7}, was also written as an extension library
for OpenFOAM, and was developed freely and openly from its inception in order to
increase community engagement and research efficiency.
